### 2022-12-05 definitions.tex
details

\problemx{(I,J)-$k$-resilience problem for WSTS}
{A state $s$ of a WSTS $(S,\rightarrow, \leq)$, an effective set $I$ (with a given basis), a set $J$.}
{$\forall s' \in J ~ (s \rightarrow^* s') \implies \exists s'' \in I ~ s' \rightarrow^{\leq k} s''$ ?\newline}

max des I,J possibles \\

J=tout \\
\problemx{general $k$-resilience problem for WSTS}
{A state $s$ of a WSTS $(S,\rightarrow, \leq)$, an effective set $I$ (with a given basis).}
{$\forall s'  ~ (s \rightarrow^* s') \implies \exists s'' \in I ~ s' \rightarrow^{\leq k} s''$ ?\newline}

@@

Analogously, we formulate the resilience problem for WSTSs.

@@

\alain{cette dernière propriété fait penser à la $t$-liveness dans un rdPetri ou n'importe quel modèle monotone..qui peut s'exprimer par $ ~ \forall s' \in S ~ (s \rightarrow^* s') \implies \exists s'' \in U_t ~ s' \rightarrow^{*} s''$ ? avec $U_t=\uparrow Pre(t)$=la clôture par le haut des préconditions pour déclencher la transition $t$}
\alain{examiner les cas I,J un élement, cloture, tout, clos par le bas et par le haut,... ...}

\begin{proposition}
coverability, liveness and reachability reduce to resilience.
\end{proposition}

@@

\begin{itemize}
\item J=tout et I=$\uparrow x$ alors resilience = couverture.
\item J=tout et I=$\downarrow x$ alors resilience = accessibilité.
\item J=tout sauf  exemple des LCS....
\end{itemize}
Let $J=\downarrow S = S$ and $I = \uparrow t$ where $t \in S$. Now resilience remains to the following property: $\forall s' \in S,  \exists s'' \in I, ~ s' \rightarrow^{*} s''$. Let us apply this property to LCS with $I = \uparrow (q,\epsilon) = (q, \Sigma^*)$. This property, $\forall s' \in S,  \exists w \in \Sigma^* $ such that $ ~ s' \rightarrow^{*} (q,w)$ expresses that there is an infinite run passing infinitely often through the control-state $q$; this LTL-property has been proved undecidable for Lossy Channel Systems (LCS) \cite{DBLP:conf/icalp/AbdullaJ94}.
Since LCS are effective WSTS with strong monotony (with an adequate semantics), we conclude.
\end{proof}

@@

citer

\begin{theorem}\cite{DBLP:conf/gg/Ozkan22,DBLP:journals/corr/abs-2108-00889}
{\sc bounded resilience} is decidable for WSTS $S$ with strong upward compatibility and the existence of an algorithm to compute a finite basis of $\uparrow post^*(S)$.
\end{theorem}

\cite{DBLP:conf/gg/Ozkan22} deduced that resilience is decidable for VASS because VASS are a finite basis of  $\uparrow post^*$ can be computed.

Context-free grammars are WSTS avec $\uparrow post^*$ calculable (because the reachability set of a pushdown machine is an effectively computable regular language) is donc resilience calculable.

We may generalyse this last result by strenghthenin to \emph{unbounded} resilience.

\begin{corollary}
{\sc resilience} is decidable for WSTS with strong upward compatibility and effective basis of $\uparrow post^*$.
\end{corollary}

\begin{proof}
The proof is essentially the same than the previous one.
\end{proof}

Il y a des pbs

\begin{proposition}
Given a class of WSTS with strong upward compatibility, there don't exist an algorithm computing an effective basis of $\uparrow post^*$.
\end{proposition}

\begin{proof}
$\uparrow post^*$ donne l'accessibilité de 0 sur les compteurs or les resert sont WSTS with strong compatibility. et l'accessibilité est undecidable.
$\uparrow post^*$ n'est pas calculable pour les LCS non plus car $\uparrow post^*= \uparrow \downarrow post^*$ n'est pas calculable ???
\end{proof}

@@

Unfortunately, resilience is undecidable for general WSTS even with strong monotony.

@@

\begin{proof} \alain{à relire/refaire}

\end{proof}

Since Reset Petri nets are also effective WSTS with strong monotony, we could think to use the undecidability of LTL for Reset Petri nets. Another proof could be a reduction of liveness to resilience and since liveness is undecidable for many models like reset PN,...resilience is also undecidable for WSTS with strong monotony.

>>>>

%
%\problemx{(I,J)-$k$-resilience problem for WSTS}
%{A state $s$ of a WSTS $(S,\rightarrow, \leq)$, an effective set $I$ (with a given basis), a set $J$.}
%{$\forall s' \in J ~ (s \rightarrow^* s') \implies \exists s'' \in I ~ s' \rightarrow^{\leq k} s''$ ?\newline}
%
%\problemx{general $k$-resilience problem for WSTS}
%{A state $s$ of a WSTS $(S,\rightarrow, \leq)$, an effective set $I$ (with a given basis).}
%{$\forall s'  ~ (s \rightarrow^* s') \implies \exists s'' \in I ~ s' \rightarrow^{\leq k} s''$ ?\newline}

@@

%Analogously, we formulate the resilience problem for WSTSs.

@@

compute the sets of $(I,J,k)$ st S is resilient.\\
max des I,J possibles \\
remark: $k$-resilienc $\leq$ bounded Resilience $\leq$ Resilience.  AVERIFIER !!!

\alain{examiner les cas I,J un élement, cloture, tout, clos par le bas et par le haut,... ...}

\begin{proposition}\label{reductions}
coverability, liveness, home state and reachability reduce to resilience.
\end{proposition}

@@

\begin{itemize}
\item $J=\uparrow s$ et $I=\uparrow x$ then this particular resilience is coverability of $x$ from $s$.
\item $J=\{s\}$ and $I=\{x\}$ this particular resilience is reachability of $x$ from $s$.
\item $J=S = \downarrow S$
\end{itemize}
Let $J=\downarrow S = S$ and $I = \uparrow t$ where $t \in S$. Now resilience remains to the following property: $\forall s' \in S,  \exists s'' \in I, ~ s' \rightarrow^{*} s''$. Let us apply this property to LCS with $I = \uparrow (q,\epsilon) = (q, \Sigma^*)$. This property, $\forall s' \in S,  \exists w \in \Sigma^* $ such that $ ~ s' \rightarrow^{*} (q,w)$ expresses that there is an infinite run passing infinitely often through the control-state $q$; this LTL-property has been proved undecidable for Lossy Channel Systems (LCS) \cite{DBLP:conf/icalp/AbdullaJ94}.
Since LCS are effective WSTS with strong monotony (with an adequate semantics), we conclude.

Since Reset Petri nets are also effective WSTS with strong monotony, we could think to use the undecidability of LTL for Reset Petri nets.

\alain{cette dernière propriété fait penser à la $t$-liveness dans un rdPetri ou n'importe quel modèle monotone..qui peut s'exprimer par $ ~ \forall s' \in S ~ (s \rightarrow^* s') \implies \exists s'' \in U_t ~ s' \rightarrow^{*} s''$ ? avec $U_t=\uparrow Pre(t)$=la clôture par le haut des préconditions pour déclencher la transition $t$}
\end{proof}

@@

Let $(X,\leq)$ be a wqo and $U$ be an upward closed subset of $X$. We say that $U$ is \emph{computable} if there is an algorithm that computes a finite basis of $U$.

\begin{theorem}\cite{DBLP:conf/gg/Ozkan22,DBLP:journals/corr/abs-2108-00889}
{\sc bounded resilience} is decidable for WSTS $S$ with strong upward compatibility and such that $\uparrow post^*(S)$ is computable.
\end{theorem}

We may immediately generalyse this last result by strenghthenin to \emph{unbounded} resilience. The proof is essentially the same than the previous one.

\begin{corollary}
{\sc resilience} is decidable for WSTS with strong upward compatibility and such that $\uparrow post^*(S)$ is computable.
\end{corollary}

However, the hypothesis that $\uparrow post^*$ is computable cannot be tested in the WSTS framework.

\begin{proposition}
There exist classes of WSTSs with strong upward compatibility for which there don't exist an algorithm computing an effective basis of $\uparrow post^*$.
\end{proposition}

\begin{proof}
Let us show that Reset VASS and LCS don't enjoy the property that $\uparrow post^*$ is computable.
Suppose that one are able to compute a finite basis of $\uparrow post^*$ for Reset VASS. Then, on ewould be able to decide whether $0$ is reachable in counter $i$ by examining if there is some vector $v$ in the basis such that $v(i)=0$. But reachability of $0$ in a counter $i$ is undecidable for Reset VASS. sont WSTS with strong compatibility. et l'accessibilité est undecidable.
$\uparrow post^*$ n'est pas calculable pour les LCS non plus car $\uparrow post^*= \uparrow \downarrow post^*$ n'est pas calculable ???
\end{proof}

@@

Unfortunately, resilience is undecidable for (general) WSTS even with strong monotony.

@@

\begin{proof} \alain{à relire/refaire}
Since reachability or liveness is undecidable for reset PN that are WSTS with strong monotony, from Proposition \ref{reductions}, we deduce that resilience is undecidable for WSTS with strong monotony.
\end{proof}

Since VASS are WSTS with strong upward compatibility and since there is an algorithm that computes a finite basis of  $\uparrow post^*(S)$, \cite{DBLP:conf/gg/Ozkan22} deduced that bounded resilience is decidable for VASS, hence resilience is decidable for VASS. \alain{resilience decidable for LCS ? LCS sont-ils strong (quelle sémantique ?) et $\uparrow post^*(S)$ is computable ?}

We may consider context-free grammars.

Context-free grammars are WSTS \cite{DBLP:journals/tcs/FinkelS01} \alain{monotony ne semble pas strong...}avec $\uparrow post^*$ calculable (because the reachability set of a pushdown machine is an effectively computable regular language) ???? resilience decidable.
